The promise of community-driven preprints in ecology and evolution

### 2024-10-16 ms/ms.qmd
Only one more comment to deal with

# 3. Following the journey of a preprint on *EcoEvoRxiv*: from submission to publication
```{r, fig-pubsummary}
#| label: fig-pubsummary
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Summary of the publication status of articles on *EcoEvoRxiv*. A) Time between uploading an preprint to *EcoEvoRxiv* and its publication as a peer-reviewed journal article. Preprints were considered those articles that were published a minimum of 2 months (60 days) after being posted to *EcoEvoRxiv*. B) Access status of published articles (both preprints and postprints) linked to *EcoEvoRxiv* preprints classified as “Open Access” or “Not Open Access”. “Unknown” status is for articles not explicitly identified as being open access or not on the Unpaywall platform. C) Sub-types of open access status of published articles on *EcoEvoRxiv*. Sub-type meanings are as follows: ‘Green’, articles published in ‘toll-access journals but achieved in an open access repository; ‘Bronze’, articles are free to read on publishers website without a license but grants no other rights and can be delayed free-to-read; ‘Hybrid’, articles are free to read upon publication with an open access license; ‘Gold’, articles published in fully open access journals. For full details on the meaning of each category see  https://support.unpaywall.org/support/solutions/articles/44001777288-what-do-the-types-of-oa-status-green-gold-hybrid-and-bronze-mean-

@@

### (a) *Science takes time, but publication could take longer*
Increased competition in science has raised the bar with respect to the amount of data required for publication [@vale2015accelerating]. This requirement is a good outcome if it results in higher-impact research that better clarifies our understanding of the natural world, but it does come at a cost for the speed of research dissemination [@vale2015accelerating]. Preprints have been proposed as a way to disseminate research more quickly as it can take a long time before results are ultimately published after formal peer review [@bourne2017ten; @vale2015accelerating]. However, data on the time to publication is needed to quantify the real benefit of preprints in this context.

@@

- Finally, keep your preprints updated. While most preprints get seamlessly connected and merged with their published version, some remain 'disconnected' as separate articles. Incorrect cross-linking by indexing platforms (e.g., Google Scholar) can create confusion and lead to frustration among authors. Thankfully, the reasons for unmatched preprints and publications are well-understood and easily rectified. They often result from a mismatch between preprint and published metadata (e.g., titles and author details). For example, nearly one-third  of articles changed their title from preprint submission to publication [`r titles[2,3]*100`% (*n* = `r titles[2,2]`)]. We found that mismatched metadata almost always contributed to preprints and published articles not being matched automatically in Google Scholar. We recommend that authors update their preprints with the publication DOI when accepted to journals, especially if their title has changed. This is very easy for authors to do on *EcoEvoRxiv* and ensures that the preprint is correctly linked to the published article and citations are appropriately merged. Adding DOIs will also create a link between the final preprint and the published paper so that, no matter the open access status of the publication, the research findings will remain openly accessible, saving costs for authors (i.e., by not having to pay open access fees) and improving the visibility and use of research.

>>>>

# 3. Following the journey of a preprint on *EcoEvoRxiv*: from submission to publication
```{r, fig-pubsummary}
#| label: fig-pubsummary
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Summary of the publication status of articles on *EcoEvoRxiv*. A) Time between uploading an preprint to *EcoEvoRxiv* and its publication as a peer-reviewed journal article. Preprints were considered those articles that were published a minimum of 2 months (60 days) after being posted to *EcoEvoRxiv*. B) Access status of published articles (both preprints and postprints) on *EcoEvoRxiv* classified as “Open Access” or “Not Open Access”. “Unknown” status is for articles not explicitly identified as being open access or not on the Unpaywall platform. C) Sub-types of open access status of published articles on *EcoEvoRxiv*. Sub-type meanings are as follows: ‘Green’, articles published in ‘toll-access journals but achieved in an open access repository; ‘Bronze’, articles are free to read on publishers website without a license but grants no other rights and can be delayed free-to-read; ‘Hybrid’, articles are free to read upon publication with an open access license; ‘Gold’, articles published in fully open access journals. For full details on the meaning of each category see  https://support.unpaywall.org/support/solutions/articles/44001777288-what-do-the-types-of-oa-status-green-gold-hybrid-and-bronze-mean-

@@

### (a) *Science takes time, but publication could take longer*
Increased competition in science has raised the bar with respect to the amount of data required for publication [@vale2015accelerating]. This requirement is a good outcome if it results in higher-impact research that better clarifies our understanding of the natural world, but it does come at a cost for the speed of research dissemination [@vale2015accelerating]. Long publication times can adversely affect EMCRs who rely on publications for job applications, promotions, and obtaining grants. Getting research out quickly can also be critical for the development of new knowledge that can re-shape research landscapes, which was important during the COVID-19 pandemic [@algaa2021development]. Preprints have been proposed as a way to disseminate research more quickly as it can take a long time before results are ultimately published after formal peer review [@bourne2017ten; @vale2015accelerating]. However, data on the time to publication is needed to quantify the real benefit of preprints in this context.

@@

- Finally, keep your preprints updated. While most preprints get seamlessly connected and merged with their published version, some remain 'disconnected' as separate articles. Incorrect cross-linking by indexing platforms (e.g., Google Scholar) can create confusion and lead to frustration among authors. The reasons for unmatched preprints and publications are well-understood and usually easily rectified. They often result from a mismatch between preprint and published metadata (e.g., titles and author details). For example, nearly one-third  of articles changed their title from preprint submission to publication [`r titles[2,3]*100`% (*n* = `r titles[2,2]`)]. We found that mismatched metadata almost always contributed to preprints and published articles not being matched automatically in Google Scholar. At times further manual merging by authors is needed to connect the preprint and published article (this can be done in Google Scholar). Regardless, we recommend that authors update their preprints with the publication DOI when accepted to journals, especially if their title has changed. This is very easy for authors to do on *EcoEvoRxiv* and should increase the chances that the preprint is correctly linked to the published article and citations are appropriately merged. Adding DOIs will also create a link between the final preprint and the published paper so that, no matter the open access status of the publication, the research findings will remain openly accessible, saving costs for authors (i.e., by not having to pay open access fees) and improving the visibility and use of research.
